---
title: "BellaBeat Leaf Case Study"
author: "Yash Naik"
date: "8/18/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# BellaBeat Product Analysis : Leaf wearable

## Deliverables

* A clear summary of the business task
* A description of all data sources used
* Documentation of any cleaning or manipulation of data
* A summary of your analysis
* Supporting visualizations and key findings
* Your top high-level content recommendations based on your analysis

## Business Task

In this project I will analyze smart device usage data in order to gain insight 
into how consumers use non-Bellabeat smart devices. I will then select one Bellabeat
product to apply these insights to, in my presentation. 

## Data Source

[FitBit Fitness Tracker Data](https://www.kaggle.com/arashnic/fitbit)
(CC0: Public Domain, dataset made available through Mobius):

This Kaggle data set contains personal fitness tracker from thirty fitbit users.
Thirty eligible Fitbit users consented to the submission of personal tracker data,
including minute-level output for physical activity, heart rate, and sleep monitoring.
It includes information about daily activity, steps, and heart rate that can be 
used to explore users’ habits.

```{r libraries, echo=TRUE}

#install.packages("tidyverse")
library(tidyverse)

#install.packages("dplyr")
library(dplyr)

#install.packages("ggplot2")
library(ggplot2)

```
## Downloaded Datasets

Leaf wearable devices is women-centric device primarily designed to track their 
activities, sleep schedules, hydration levels, hygene, etc.
I will be analyzing Fitbit smart wearable device which provides tracking of some
of the similar activities if not all. So, I have included these below mentioned 
datasets for my analysis.

* dailyActivity_merged
* dailySteps_merged
* minuteSleep_merged
* sleepDay_merged

```{r dataset, echo=FALSE}
sleep_log<- read_csv("~/Downloads/GOOGLE_DATA_ANALYTICS/capstone/fitbit_dataset/sleepDay_merged.csv")
activity_log <- read_csv("~/Downloads/GOOGLE_DATA_ANALYTICS/capstone/fitbit_dataset/dailyActivity_merged.csv")
sleep_minutes <- read_csv("~/Downloads/GOOGLE_DATA_ANALYTICS/capstone/fitbit_dataset/minuteSleep_merged.csv")
steps_log <- read_csv("~/Downloads/GOOGLE_DATA_ANALYTICS/capstone/fitbit_dataset/dailySteps_merged.csv")
```

#### Explore the datasets
```{r explore, echo = TRUE}

head(sleep_log)
head(activity_log)
head(steps_log)
head(sleep_minutes)
```

Checking the maximum and minimum distance values in activity dataset to look for 
potential outliers


```{r outliers, echo=TRUE}
max(activity_log$TrackerDistance)
min(activity_log$TrackerDistance)

max(sleep_log$TotalMinutesAsleep)
min(sleep_log$TotalMinutesAsleep)
```

We can observe that minimum value for TrackerDistance is 0 which is an indication
to examine further.

```{r zero_vals, echo=TRUE}

activity_log %>% count(activity_log$TrackerDistance == 0)

```
We can see that out of 940 results 78 are 0 which accounts for roughly 5% missing
values from daily activities, so it will be good idea to filter out these missing
values.

```{r filtered, echo=TRUE}

activity_filtered <- activity_log %>% 
  filter(activity_log$TrackerDistance != 0)

head(activity_filtered)
```

Next, we will calculate summary statistics of the filtered activity data and 
create different levels for activity.



```{r summary_activity, echo=TRUE}

summary_activity <- activity_filtered %>% 
  group_by(Id) %>% 
  summarize(total_dist = sum(TrackerDistance), avg_dist = mean(TrackerDistance),
            num_dist = length(TrackerDistance))

summary_activity$distance_level <- case_when(
  summary_activity$avg_dist >= 8 ~ "Highly Active",
  summary_activity$avg_dist >= 4 ~ "Moderately Active",
  summary_activity$avg_dist >= 1 ~ "lowly Actively",
  summary_activity$avg_dist < 1 ~ "inactive"
)

head(summary_activity)

```
Creating a scatter plot to verify if there is any positive correlation between 
tracking distance and average distance.

```{r plot1, echo=TRUE}

ggplot(data=summary_activity) + geom_point(aes(x = avg_dist, y = num_dist, 
                                               color= distance_level)) + 
  facet_wrap(~distance_level)

```
From the plot above, we can observe a *little* *positive* *correlation* between
tracking entries and average distance covered. All the users who are highly active
have logged their entries more than 20 times. Similarly, moderately active users,
have comparatively more tracking entries than lowly_active users.

Now, we can extend our analysis to dig deeper and check if users who are highly
active, have slept for longer.

```{r sleep_track, echo= TRUE}

summary_sleep <- sleep_log %>% 
  group_by(Id) %>% 
  summarize(total_sleep_min = sum(TotalMinutesAsleep), average_sleep_min = mean(TotalMinutesAsleep), num_sleep_entries = length(TotalMinutesAsleep))

summary_sleep$sleep_levels <- case_when(
  summary_sleep$average_sleep_min >= 480 ~ "Well Rested",
  summary_sleep$average_sleep_min >= 360 ~ "Moderately Rested",
  summary_sleep$average_sleep_min >= 240 ~ "Poorly Rested",
  summary_sleep$average_sleep_min < 240 ~ "Not Rested"
)
  
head(summary_sleep)
```
Now, we will again create a scatter plot to check for any correlation between
average sleep and number of sleep entries.

```{r sleep_plot, echo=TRUE}

ggplot(data = summary_sleep) + geom_smooth(aes(x = num_sleep_entries, y = average_sleep_min))

```
Yes, we can observe a *positive* *correlation* between users who have more sleep
entries and the number of hours they have slept.

### To analyse a pattern between activity and sleep, we can combine the two summary dataframes.

Combining *summary_activity* and *summary_sleep* to create a single dataframe.

```{r activity_sleep,echo=TRUE}

activity_sleep <- merge(summary_activity, summary_sleep, by = 'Id')

head(activity_sleep)

```
We will again visualize to check if there is any correlation between users who
are highly active and who have slept for longer duration.

## Visualization

```{r plot2, echo=TRUE}

ggplot(data= activity_sleep) + geom_bar(aes(x=sleep_levels, fill = distance_level)) +
  labs(title = "Sleep levels VS Distance Levels")

ggplot(data= activity_sleep) + geom_bar(aes(x=distance_level, fill = sleep_levels)) +
  labs(title = "Sleep levels VS Distance Levels")

ggplot(data = activity_sleep) +
   geom_line(aes(x=avg_dist, average_sleep_min))+
   geom_smooth(aes(x=avg_dist, average_sleep_min))+
   labs(title='Average Distance Tracked VS Average Sleep Time')

ggplot(data=activity_sleep)+
  geom_line(aes(x=average_sleep_min, avg_dist))+
  geom_smooth(aes(x=average_sleep_min,y=avg_dist))+
  labs(title='Average Sleep Time VS Average Distance Tracked')
```
Initially, I thought there would be a strong positive correlation between average
distance tracking and sleeping i.e. A person who is very actively should have 
rested for a longer duration than someone who is less active on an average. Looking 
at these trends, I am surprised that my assumption has been refuted as after 
visualizing these trends, it appears that there is *little* to *no* correlation 
between the two things.

```{r, echo=TRUE}
ggplot(data = activity_sleep) + geom_line(aes(x = average_sleep_min, y = avg_dist))+
  labs(title = "Average Distance tracked VS Sleep levels")+
  facet_wrap(~distance_level)
```
* Thus, we can observe that a majority of people who tracked as average distance
of > 7 Km rested for less than 400 minutes (6 hours, 30 mins)

* On Contrary, people who were moderately active, slept for longer than 400 minutes.


## Summary

* This analysis was performed on a Fitbit wearable smart tracking device. The data 
collected was based on a small group of 30 people in 2016 which was more than 5
years ago.

* While cleaning the data, I found that almost *5%* of the data from the sleep log 
was *missing*, which could mean different things like - some users did not wear the 
device while sleeping or the device got discharged at night or there was some 
technical issue in the device which resulted in these *0* values.

* Contrary to my initial assumption (people who are more active should sleep for 
longer duration), I observed that there is *no correlation* between increased 
distance tracking and sleeping for longer duration. 

* Final results show that people who were moderately active diring the day i.e.
people who tracked a distance between 4 Km to 8 Km had an average sleep of 6 hours 
to 8 hours and beyond. 


## Recommendation

After analyzing the fitbit smart device in order to have its comparision with 
Bellabeat's Leaf, I can conclude with the following recommedations-

* Focus on advertisements to market the theory that tracking increased distances
using Leaf device may result in better/relaxed sleep. 

* We should also consider why there were 5% missing values in Fitbit devices and 
avoid having those similar trends in Leaf by focusing on R & D as well.

** Increasing battery capacity to avoid discharge during the nights or decrease
battery usage through *low power mode* etc. 


